---
title: "Import CSV from ConfoMap ISO25178 - use-wear archaeology & experiments"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output: word_document
---


```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment = NA, message = FALSE, indent = "", error = TRUE)

```

---


# Goal of the script
This script formats the output resulting from applying a template computing ISO 25178 parameters in ConfoMap.
The script will:

1. Read in the original Rbin-files
2. Format the data in order to make the three data sets fitting  
3. Write an XLSX-file and save an R object ready for further analysis in R 

```{r}
dir_in <- "analysis/raw_data"
dir_out <- "analysis/derived_data/"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(openxlsx)
library(tools)
library(R.utils)
library(chron)
library(tidyverse)

```


---


# Get names, path and information of all files 
```{r}
data_files <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
md5_in <- md5sum(data_files)
info_in <- data.frame(files = basename(names(md5_in)), checksum = md5_in, 
                  row.names = NULL)

```

The checksum (MD5 hashes) of the imported files are:  
```{r, echo = FALSE}
info_in

```


---


# Read in original CSV-files
```{r}
# data from 'aVSn' experiment, tool function experiment and archaeology 
AvsN_imp <- loadObject(data_files[1])
str(AvsN_imp)
TFE_imp <- loadObject(data_files[2])
str(TFE_imp)
Arch_imp <- loadObject(data_files[3])
str(Arch_imp)

```


---

# Add columns with further information 
```{r}

location <- substr(AvsN_imp["Location"][,1], 1, 1)
sublocation <- substr(AvsN_imp["Location"][,1], 2, 2)

AvsN_imp$Location = location
AvsN_imp$Sublocation = sublocation


location <- substr(TFE_imp["Location"][,1], 1, 1)
sublocation <- substr(TFE_imp["Location"][,1], 2, 2)

TFE_imp$Location = location
TFE_imp$Sublocation = sublocation

# adds the name of the experiment as "site"
AvsN_imp[grep("LYDIT4-", AvsN_imp[["Sample"]]), "Site"] <- "aVSn"
AvsN_imp[grep("FLT4-", AvsN_imp[["Sample"]]), "Site"] <- "aVSn"
TFE_imp[grep("LYDIT5-", TFE_imp[["Sample"]]), "Site"] <- "TFE"
TFE_imp[grep("FLT8-", TFE_imp[["Sample"]]), "Site"] <- "TFE"

AvsN_imp[["Site"]] <- factor(AvsN_imp[["Site"]])
TFE_imp[["Site"]] <- factor(TFE_imp[["Site"]])

# adds the tool type 
AvsN_imp[grep("aVSn", AvsN_imp[["Site"]]), "Tool.type"] <- "Standard sample"
TFE_imp[grep("TFE", TFE_imp[["Site"]]), "Tool.type"] <- "Standard sample"

AvsN_imp[["Tool.type"]] <- factor(AvsN_imp[["Tool.type"]])
TFE_imp[["Tool.type"]] <- factor(TFE_imp[["Tool.type"]])

# adds a column for "use-wear type" based on cycle
AvsN_imp$Usewear.type <- AvsN_imp$Cycle
TFE_imp$Usewear.type <- TFE_imp$Cycle

# adds the contact material 
Arch_imp[grep("lydite", Arch_imp[["Raw.material"]]), "Contact.material"] <- "unknown"
Arch_imp[grep("flint", Arch_imp[["Raw.material"]]), "Contact.material"] <- "unknown"

# adds the task 
Arch_imp[grep("lydite", Arch_imp[["Raw.material"]]), "Task"] <- "unknown"
Arch_imp[grep("flint", Arch_imp[["Raw.material"]]), "Task"] <- "unknown"
AvsN_imp[grep("aVSn", AvsN_imp[["Site"]]), "Task"] <- "cutting"

# adds the edge angle 
Arch_imp[grep("lydite", Arch_imp[["Raw.material"]]), "Edge.angle"] <- "unknown"
Arch_imp[grep("flint", Arch_imp[["Raw.material"]]), "Edge.angle"] <- "unknown"
AvsN_imp[grep("aVSn", AvsN_imp[["Site"]]), "Edge.angle"] <- "60°"

```


# Keeps only interesting columns and orders them  
```{r}
# deletes non-important columns 
AvsN_imp$Cycle <- NULL
TFE_imp$Cycle <- NULL

# orders the columns in an identical way
AvsN_imp <- AvsN_imp[c(1, 56:57, 6, 2, 55, 3:4, 58, 7, 59:60, 5, 8:54)]
TFE_imp <- TFE_imp[c(1, 58:59, 6, 2, 57, 3:4, 60,7:9, 5, 10:56)]
Arch_imp <- Arch_imp[c(1:9, 58:60, 10:57)]

```


# Merges the three datasets
```{r}
# check pairwise if the three lines of headers are identical among the datasets
# merges the data based on the three lines of headers while they get only 
# used in the first file 

comp <- all(sapply(list(colnames(AvsN_imp), colnames(TFE_imp)), 
                   FUN = identical, colnames(Arch_imp)))
if (comp == TRUE) {
  merged_data <- full_join(AvsN_imp, TFE_imp) %>% full_join(Arch_imp)
} else {
  stop("The headers are not identical among the datasets")
}
str(merged_data)

```


---


# Save data
## Format name of output file
```{r}
file_out <- "Use-wear_all"

```
The files will be saved as "`r paste0("~/", file_out, ".[ext]")`".


## Write to XLSX
```{r}
write.xlsx(list(data = merged_data), 
           file = paste0(dir_out, file_out, ".xlsx"))

```


## Save R object
```{r}
saveObject(merged_data, file = paste0(dir_out, file_out, ".Rbin"))

```

---

# sessionInfo() and RStudio version
```{r}
sessionInfo()

```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---


END OF SCRIPT







