---
title: "Plots - use-wear archaeology & experiments"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", 
  knit_root_dir = rprojroot::find_rstudio_root_file()) })
---

```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)
````


# Goal of the script
This script plots all variables to see which ones should be used for further analysis.  
Scatterplot of each variable will be plotted.   



```{r}
dir_in <- "analysis/derived_data/"
dir_out <- "analysis/plots"
```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.
The knit directory for this script is the project directory.

---


# Load packages
```{r}
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
library(patchwork)
library(doBy)
library(ggrepel)
library(openxlsx)
library(wesanderson)
library(ggfortify)

```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)
```


The checksum (MD5 hashes) of the imported file is:  
```{r, echo = FALSE}
info_in
```


# Load data into R object
```{r}
imp_data <- loadObject(data_file)
str(imp_data)
```

The imported file is: "`r paste0("~/", data_file)`"  

---


# Prepare variables
## Define numeric variables
```{r}
num.var <- 27:length(imp_data)
```

The following variables will be used: 

```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep="")
```

---


# Plot each of the selected numeric variables 
## Boxplot of all the variables combined with the tool type 
```{r}

arch <- filter(imp_data, Site == "Balve" | Site == "Buhlen" | Site == "Ramioul") 
exp <- filter(imp_data, Site == "aVSn" | Site == "TFE")

# converts the data into factor 
arch[["Tool.type"]] <- factor(arch[["Tool.type"]])
exp[["Tool.type"]] <- factor(exp[["Tool.type"]])

# colours archaeological data
custom.col2 <- data.frame(type = levels(arch$Tool.typ), 
                         col = c( "#0B775E", "#E1BD6D", "#35274A" ,"#F2300F")) 
arch$col <- custom.col2[arch$Tool.typ %in% levels(arch[["Tool.type"]]), "col"]

# colours experimental data 
# Darjeeling1 = c("#FF0000", "#00A08A", "#F2AD00", "#F98400", "#5BBCD6")
custom.col6 <- data.frame(type = levels(exp$Tool.typ), col = "#F98400") 
exp$col <- custom.col6[exp$Tool.typ %in% levels(exp[["Tool.type"]]), "col"]



# plot tool type 
for (i in num.var){
  
  # gets the min/max range of the data set
range_var <- range(imp_data[[i]]) 

  
  p_arch <- ggplot(data = arch, aes_string(x = "Usewear.type", y = names(imp_data)[i],
                                           fill = "Tool.type")) +
        geom_boxplot(outlier.size = 0.7, lwd = 0.2) +
        theme_classic() +
        coord_cartesian(ylim = range_var) + 
        labs( x = "Use-wear type", title = " ") +
        labs(y = gsub("\\.", " ", names(imp_data)[i])) +
        labs(fill = "Tool type") + 
        scale_fill_manual(values = arch$col)
        
  
  p_exp <- ggplot(data = exp, aes_string(x = "Usewear.type", y = names(imp_data)[i],
                                           fill = "Tool.type")) +
        geom_boxplot(outlier.size = 0.7, lwd = 0.2) +
        theme_classic() +
        coord_cartesian(ylim = range_var) + 
        labs( x = "Use-wear type", title = " ") +
        labs(y = gsub("\\.", " ", names(imp_data)[i])) +
        labs(fill = "Tool type") + 
        scale_fill_manual(values = exp$col)
  

  # combines the plots 
  p <- p_arch + p_exp + plot_layout(width = c(4, 1), guides = 'collect')
 
  print(p)
   
  # saves the plots 
  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_boxplot_arch.exp_", 
	                   names(imp_data)[i], ".pdf")
	ggsave(filename = file_out, plot = p, path = dir_out, device = "pdf", width = 250, height = 170, units = "mm")
      
}  

```



## Principal component analysis 
```{r}
# removes rows with na values 
data_pca <- na.omit(imp_data)
# uses for the PCA only selected variables: Sq, SSK, Vmc, Isotropy, Mean density of furrows, Asfc, HAsfc9 
imp_data.pca <- prcomp(data_pca[, c(27:28, 45, 50, 54, 57, 59)], scale. = TRUE) 
# converts the data into factor 
data_pca[["Usewear.type"]] <- factor(data_pca[["Usewear.type"]])


custom.col <- data.frame(type = levels(data_pca$Usewear.typ), 
                         col = c("#FD6467", "#999999", "#52854c", "#c3d7a4", "#00A08A", "#487bb6", "#9a0f0f", "#fdbf6f",
                                 "#d16103", "#ffdb6d", "#985633", "#134680", "#05100c")) 
data_pca$col <- custom.col[data_pca$Usewear.typ, "col"]


# Using ggfortify
PCA<- autoplot(imp_data.pca, data = data_pca, colour = "Usewear.type", size = 2,
             loadings = TRUE, loadings.colour = "black", loadings.label = TRUE, loadings.label.colour = "black", 
             loadings.label.size  = 4, loadings.label.repel = TRUE,  
             frame = TRUE, frame.type = "convex", frame.colour = "Usewear.type", frame.alpha = 0) + 
             theme_classic() +
             scale_colour_manual(values = custom.col$col)
         
print(PCA)

# saves the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_PCA_all.data", ".pdf")
ggsave(filename = file_out, plot = PCA, path = dir_out, device = "pdf")
  
```


## (facet plot = 1 plot for flint, 1 plot for lydite)
```{r}
# selects only the data from after 2000 strokes on the bone plate (experimental data) 
exp2000 <- filter(exp, Usewear.type == "2000" & Contact.material == "bone plate")

for (i in num.var){
  #plot
  range_var <- range(exp2000[[i]]) # gets the min/max range of the data set 
  
	p <- ggplot(data = exp2000, aes_string(x="Task", y = names(exp2000)[i], colour="Edge.angle")) + 
	     # avoids overplotting
	     scale_color_manual(values=c("#446455","#999999", "#E69F00")) +
	     geom_jitter(size = 3,  position = position_jitter(width = 0.35, seed = 1)) + 
	     coord_cartesian(ylim = range_var) + 
	     theme_classic() + 
	     # removes the "." between "Edge.angle" in the legend 
	     labs(colour = gsub("\\.", " ", "Edge.angle")) + 
	     # removes the "." between the variable names
	     labs(y = gsub("\\.", " ", names(exp2000)[i])) + 
       # combines the flint and the lydite plot into one   
	     facet_wrap(~Raw.material) 
	     
	
	print(p)


	#save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_exp_2000_plot_", 
	                   names(imp_data)[i], ".pdf")
	ggsave(filename = file_out, plot = p, path = dir_out, device = "pdf")
}


```



The files will be saved as "`r paste0("~/", dir_out, ".[ext]")`".

---


# Show plot files information
```{r}
info_out <- list.files(path = dir_out, pattern = "\\.pdf$", 
                       full.names = TRUE) %>% 
            md5sum()
```


The checksum (MD5 hashes) of the exported files are:  
```{r, echo = FALSE}
info_out
```



---

# sessionInfo() and RStudio version
```{r}
sessionInfo()
```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---

END OF SCRIPT

