---
title: "Plots - use-wear archaeology & experiments"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r Knitr Options, include=FALSE}
knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)

```


# Goal of the script
This script combines the results from the quantitative use-wear analysis performed on archaeological samples as well as on standard samples used during two experiments (aVSn and TFE)


```{r}
dir_in <- "analysis/derived_data/"
dir_out <- "analysis/plots"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.
The knit directory for this script is the project directory.

---

# Load packages
```{r Libraries}
pack_to_load <- c("tools", "openxlsx", "ggplot2", "doBy", "R.utils", "tidyverse", "patchwork", "ggrepel", "ggfortify")
sapply(pack_to_load, library, character.only = TRUE, logical.return = TRUE)

```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)

```


# Load data into R object
```{r}
imp_data <- loadObject(data_file)
str(imp_data)

```

The imported file is: "`r paste0("~/", data_file)`"  

---


# Prepare variables
## Define numeric variables
```{r}
num.var <- 27:length(imp_data)

```

The following variables will be used: 

```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep="")

```

---


# Plot each of the selected numeric variables 
## Boxplot of all the variables combined with the artefact category (without outliers) 
```{r}
# exclude the outliers (from both data sets)
# add the indices as row numbers 
imp_data <- imp_data %>% mutate(id = row_number())
imp_data2 <- imp_data[-c(19, 20, 55, 59, 62, 70, 139, 147, 199, 200), ]


arch <- filter(imp_data2, Site == "Balve" | Site == "Buhlen" | Site == "Ramioul") 
exp <- filter(imp_data2, Site == "aVSn" | Site == "TFE")

# convert the data into factor 
arch[["Tool.type"]] <- factor(arch[["Tool.type"]])
exp[["Tool.type"]] <- factor(exp[["Tool.type"]])

# colours archaeological data
custom.col2 <- data.frame(type = levels(arch$Tool.typ), 
                         col = c( "#0B775E", "#E1BD6D", "#F2300F", "#35274A")) 
arch$col <- custom.col2[arch$Tool.typ %in% levels(arch[["Tool.type"]]), "col"]

# colours experimental data 
# Darjeeling1 = c("#FF0000", "#00A08A", "#F2AD00", "#F98400", "#5BBCD6")
custom.col6 <- data.frame(type = levels(exp$Tool.typ), col = "#F98400") 
exp$col <- custom.col6[exp$Tool.typ %in% levels(exp[["Tool.type"]]), "col"]

# get new order 
exp$Usewear.type <- factor(exp$Usewear.type, levels=c("before", "2000"))


# plot tool type 
for (i in num.var){
  
  # get the min/max range of the data set
range_var <- range(imp_data2[[i]]) 

  
  p_arch <- ggplot(data = arch, aes_string(x = "Usewear.type", y = names(imp_data)[i],
                                        fill = "Tool.type")) +
            geom_boxplot(outlier.size = 0.7, lwd = 0.2) +
            theme_classic() +
            coord_cartesian(ylim = range_var) + 
            labs( x = "use-wear type", title = " ") +
            labs(y = gsub("\\.", " ", names(imp_data)[i])) +
            labs(fill = "artefact category") + 
            scale_fill_manual(values = arch$col)
        
  
  p_exp <- ggplot(data = exp, aes_string(x = "Usewear.type", y = names(imp_data)[i],
                                      fill = "Tool.type")) +
           geom_boxplot(outlier.size = 0.7, lwd = 0.2) +
           theme_classic() +
           coord_cartesian(ylim = range_var) + 
           labs( x = "use-wear type", title = " ") +
           labs(y = gsub("\\.", " ", names(imp_data)[i])) +
           labs(fill = "artefact category") + 
           scale_fill_manual(values = exp$col)
  

  # combine the plots 
  p <- p_arch + p_exp + plot_layout(width = c(4, 1), guides = 'collect')
 
   # save the plots 
  file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_boxplot_arch.exp_", 
	                   names(imp_data)[i], ".pdf")
	ggsave(filename = file_out, plot = p, path = dir_out, device = "pdf", 
	         width = 250, height = 170, units = "mm")
      
}  

```


## Principal component analysis (without outliers)
```{r}

# PCA use-wear types
# remove rows with na values 
data_pca <- na.omit(imp_data2)
# use for the PCA only selected variables: Sq, SSK, Vmc, Isotropy, Mean density of 
# furrows,  Asfc, HAsfc9 
imp_data.pca <- prcomp(data_pca[, c(27:28, 45, 50, 54, 57, 59)], scale. = TRUE) 
# convert the data into factor 
data_pca[["Usewear.type"]] <- factor(data_pca[["Usewear.type"]])


custom.col1 <- data.frame(type = levels(data_pca$Usewear.typ), 
                           col = c("#FD6467", "#999999", "#52854c", "#c3d7a4", "#00A08A",
                                   "#487bb6", "#9a0f0f", "#fdbf6f", "#d16103", "#ffdb6d",
                                   "#985633", "#134680", "#05100c")) 
data_pca$col <- custom.col1[data_pca$Usewear.typ, "col"]


# using ggfortify
PCA<- autoplot(imp_data.pca, data = data_pca, colour = "Usewear.type", size = 2,
                         loadings = TRUE, loadings.colour = "black", loadings.label = TRUE,                          loadings.label.colour = "black", 
                         loadings.label.size  = 4, loadings.label.repel = TRUE,  
                         frame = TRUE, frame.type = "convex", frame.colour = "Usewear.type",
                         frame.alpha = 0) + 
       theme_classic() +
       scale_colour_manual(values = custom.col1$col)

# save the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_PCA_all.data", ".pdf")
ggsave(filename = file_out, plot = PCA, path = dir_out, device = "pdf")

```


## (facet plot = 1 plot for flint, 1 plot for lydite)
```{r}
# select only the data from after 2000 strokes on the bone plate (experimental data) 
exp2000 <- filter(exp, Usewear.type == "2000" & Contact.material == "bone plate")

for (i in num.var){
  # plot
  range_var <- range(exp2000[[i]]) # get the min/max range of the data set 
  
	p <- ggplot(data = exp2000, aes_string(x = "Task", y = names(exp2000)[i],
	          colour = "Edge.angle")) + 
	     # avoid overplotting
	     scale_color_manual(values=c("#046C9A","#FAD510", "#CB2314")) +
	     geom_jitter(size = 3,  position = position_jitter(width = 0.35, seed = 1)) + 
	     coord_cartesian(ylim = range_var) + 
	     theme_classic() + 
	     # remove the "." between "Edge.angle" in the legend 
	     labs(colour = gsub("\\.", " ", "Edge.angle")) + 
	     # remove the "." between the variable names
	     labs(y = gsub("\\.", " ", names(exp2000)[i])) + 
       # combine the flint and the lydite plot into one   
	     facet_wrap(~Raw.material) 


	#save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_exp_2000_plot_", 
	                   names(imp_data)[i], ".pdf")
	ggsave(filename = file_out, plot = p, path = dir_out, device = "pdf")
}

```

The files will be saved as "`r paste0("~/", dir_out, ".[ext]")`".

---


# sessionInfo() and RStudio version
```{r}
sessionInfo()

```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


# Cite R packages used
```{r Citation}
for (i in pack_to_load) print(citation(i), bibtex = FALSE)

```


---

END OF SCRIPT

